---
title: "Tidyverse Solo Exercise"
output: html_notebook
---
1. Read in `districts.csv` into a tibble named `districts`.

---
```{r}
library(tidyverse)
getwd()
districts <- read.csv("../data/districts.csv")


```
---
2. Notice that the first row corresponds to the whole State of Tennessee. Remove this row and save the result back to `districts`.
---
```{r}
districts <- districts[-1, ]
```
---
3. How many districts have a proficiency rate of at least 80% for both alg_1 and eng_1?

---
```{r}
cleaned_district <- districts %>%
  filter(!is.na(alg_1), !is.na(eng_1))
alg1_eng1_80_filter <- filter(cleaned_district, alg_1 >= 80, eng_1 >= 80)
sorted_district <- alg1_eng1_80_filter %>%
  mutate(performance_score = (alg_1 + eng_1)/2)
districts_rank <- sorted_district %>%
  arrange(desc(performance_score))
top_districts <- districts_rank$system_name
top_districts_string <- paste(top_districts, collapse = ",")
row_count <- nrow(alg1_eng1_80_filter)
cat("Number of districts that have a proficiency rate of at least 80% for both Algebra and English is", row_count, ".\n", "The top 13 districts are ranked based on their performance, starting with the best-performing district are", top_districts_string, ".")
```
4. How many districts have a proficiency rate less than 50% for either alg_1 or eng_1?


```{r}
bad_performance_dist <- filter(cleaned_district, alg_1 < 50 | eng_1 <50)
dist_count2 <- nrow(bad_performance_dist)
underperform_dist_name <- bad_performance_dist$system_name
name_list <- paste(underperform_dist_name, collapse = ",")
cat("Number of districts that have a proficiency rate of alg_1 or eng_1 less than 50% is", dist_count2, ".", "The districts requiring improvement in either Algebra scores or English scores are:", name_list, ".")
```
5. Which district has the lowest graduation rate?
```{r}
cleaned_data <- districts %>%
  filter(!is.na(enrollment), !is.na(grad))
cleaned_data <- cleaned_data %>%
  mutate(graduation_rate = grad/enrollment * 100 )
lowest_grad_dist <- cleaned_data %>%
  arrange(graduation_rate) 
lowest_grad_dist_name <- lowest_grad_dist$system_name[1]
cat("The district with the lowest graduation race is", lowest_grad_dist_name, ".")
```
6. Within the Mid Cumberland region, which district has the highest ACT composite?
```{r}
Mid_Cum_filter <- districts %>%
  filter(region == "Mid Cumberland", !is.na(act_composite)) %>%
  arrange(desc(act_composite))
higest_ACT_dist <- Mid_Cum_filter$system_name[1]
cat("The district witht the higest ACT composite in the Mid Cumberland region is", higest_ACT_dist, ".")

```
7. Create a scatter plot to compare alg_1 proficiency rates to alg_2 rates. What do you notice? Facet this plot by region. Does anything stand out when you facet the plots?
```{r}
select_data <- districts %>%
  filter(!is.na(alg_1), !is.na(alg_2), !is.na(region))
ggplot(select_data,aes(x= alg_1, y = alg_2)) +
  geom_point(color = "red") +
  facet_wrap(~region, ncol = 2) +
  labs(title = "Comparison of Algebra 1 and Algebra 2 proficiency Rates by region", x = "Algebra 1 Proficiency Rate", y = "Algebra 2 Proficiency Rate")+
  theme(plot.title = element_text(hjust = 0.5))

```
---
Summary
1. There is  a positive relationship between Algebra 1 and Algebra 2 proficiency rates across regions
2. Some regions like Southwest/Memphis and Mid Cumberland shoo greater variation among districts within those region. In contrast, Northwest and First TN have more uniform performance among their districts
3. Some regions with point representing low in both Algebra 1 and Algebra 2 will need a plan to improve proficiency rates.
---

8. When creating this bar chart you may have noticed that some districts have missing enrollment values. For how many districts is this the case?
```{r}
count_missing <- sum(is.na(districts$enrollment))
cat("Number os districts that have missing enrollment values is", count_missing, ".")
```
9. A lot of rows are missing additional values. Which district has the largest number of missing values (across all variables)? Hint: you might want to look at rowwise and c_across to answer this question.
```{r}
rowwise_data <- districts %>%
  rowwise() %>%
  mutate(all_miss_count = sum(is.na(c_across(alg_1:dropout))))
max_na_row <- rowwise_data %>%
  arrange(desc(all_miss_count))
dist_name <- max_na_row$system_name[1]
cat("The district with the largest number of missing values is", dist_name, ".")
```
10. What is the mean graduation rate across all districts? What might be wrong with using just the regular mean to assess average graduation rates?
```{r}
cleaned_grad_column <- districts %>%
  filter(!is.na(grad))
mean_all_dist_grad <- mean(cleaned_grad_column$grad)
cat("The mean graduation rate across all districts is", mean_all_dist_grad, ".")
ggplot(cleaned_grad_column, aes(x = system_name, y = grad)) +
  geom_point(color = "red") +
  labs(
    title = "Distribution of Graduation Rate Across All Districts",
    x = "District Name",
    y = "Graduation Rate") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 6), plot.title = element_text(hjust =0.5)) +
  scale_x_discrete(breaks = cleaned_grad_column$system_name[seq(1, nrow(cleaned_grad_column), by = 2)])
```
---
Using the mean graduation rate across all districts to assess average graduation rates in Tennessee will misrepresent the overall performance of the state's school system. It does not capture how many districts are performing above or below the mean adn can overlook many districts with low graduation rates.
---
11. Redo the previous question but use a weighted average (`weighted.mean`) graduation across all districts, weighing by enrollment. How much does this change your answer? Can you explain using the data the reason for the big change from using the mean?
```{r}
cleaned_grad_column <- cleaned_grad_column %>%
  filter(!is.na(enrollment))
weighted_mean_graduate <- sum(cleaned_grad_column$grad*cleaned_grad_column$enrollment)/sum(cleaned_grad_column$enrollment)
cat("The weighted average graduation across all districts is", weighted_mean_graduate, ".", "The weighted gradutaion rate is lower in districts that enrolled high number of students")
```
12. Find the unweighted and weighted average graduation rate by district.
```{r}
unweighted_grad <- mean(cleaned_grad_column$grad)
weighted_grad_by_black <- sum(cleaned_grad_column$grad * cleaned_grad_column$black)/sum(cleaned_grad_column$black)
weigted_grad_by_his <- sum(cleaned_grad_column$grad*cleaned_grad_column$hispanic)/sum(cleaned_grad_column$hispanic)
weight_grad_by_native <- sum(cleaned_grad_column$grad*cleaned_grad_column$native)/sum(cleaned_grad_column$native)
weight_grad_by_exp <- sum(cleaned_grad_column$grad*cleaned_grad_column$expenditures)/sum(cleaned_grad_column$expenditures)
```

**Continued Exploration and Practice**

13. Read in the school-level testing data for 2014, available https://docs.google.com/spreadsheets/d/1i6s9wPs5lg9fFqadCfW9qccPwuKSIu2t/edit?usp=sharing&ouid=102586165284403190694&rtpof=true&sd=true. You might find the readxl library useful for this task. If you use this library, be sure to look at the `na` argument for the `read_excel` function.
```{r}
library(readxl)
school_test <-
```


To answer the following questions, use "All Students" for the subgroup. 

14. How many schools have at least 20 percent of students below bsc for Algebra I? Which districts do these schools belong to?

15. How many schools have at least 20 percent of students below bsc for _both_ Algebra I and English I?

16. Which grade has the highest pct_adv for Algebra I? Plot the average pct_adv per grade level as a bar chart. Make sure that the bars are ordered by grade level.

17. Find the correlation between pct_adv for Algebra I and pct_adv for Algebra II by school. Create a scatterplot showing Algebra II scores vs. Algebra I scores by school.

18. Find all schools in Rutherford County that have "High School" in their name. For these schools, create a chart (your choice) showing the differences in pct_below_bsc, pct_bsc, pct_prof, and pct_adv for Algebra I when looking across all subgroups and grades.

19. Create a function which allows you to select a system_name and which creates a plot to show the differences in pct_below_bsc, pct_bsc, pct_prof, and pct_adv for Algebra I when looking across all subgroups and grades for all schools with "High School" in their names within that system.

